---
title: "extended_dataset"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(stringdist)
library(ggrepel)
library(recipes)
library(glue)
```

```{r}
data <- read.csv("~/Desktop/dougscore-dataset/DougScore.csv")
data_2 <- read.csv("~/Desktop/dougscore-dataset/ds_updated.csv")
```

```{r}
api_key <- Sys.getenv("OPENAI_API_KEY")

if (api_key == "") {
  stop("OpenAI API key not found. Make sure it's in your .Renviron file.")
}
```


```{r}
dougscore <- data_2 %>% 
  mutate(
    weekend_score = styling + acceleration + handling + fun_factor + cool_factor,
    daily_score = features + comfort + quality + practicality + value,
    total_score = weekend_score + daily_score, 
    car_id = paste(year, make, model)
  )
```


#Functions

##Get Top Cars
```{r}
get_top_cars <- function(df, metric, n = 5) {
  if (!metric %in% colnames(df)) {
    stop(paste("Column", metric, "not found in dataframe."))
  }

  df %>%
    arrange(desc(.data[[metric]])) %>%
    slice_head(n = n)
}
```

###Test: Get Top Cars
```{r}
get_top_cars(dougscore, metric = "msrp", n = 10)
```

##Search Cars
```{r}
search_cars <- function(df, query, metric = NULL, max_results = 10) {
  results <- search_cars(df, query, max_results = max_results)

  # If a metric is supplied, return just the core info + metric
  if (!is.null(metric)) {
    if (!metric %in% colnames(df)) {
      stop(paste("Metric", metric, "not found in dataset."))
    }

    return(results %>%
             mutate(car_id = paste(year, make, model)) %>%
             select(car_id, !!sym(metric)))
  }

  # Otherwise, just return the full search result
  return(results)
}
```

###Test: Search Cars
```{r}
search_cars(dougscore, "gt-r")
search_cars(dougscore, "ferrari", max_results = 10)
search_cars(dougscore, "1994", max_results = 5)
```


##Filter By
```{r}
filter_by <- function(df, column, query, max_results = 10) {
  column <- rlang::ensym(column)
  query <- tolower(query)

  df %>%
    filter(str_detect(tolower(!!column), fixed(query))) %>%
    slice_head(n = max_results)
}
```

###Test: Filter By
```{r}
gtr <- filter_by(dougscore, model, "gtr")
head(gtr)
porsche <- filter_by(dougscore, make, "porsche")
head(porsche)
year_1994 <- filter_by(dougscore, year, "1994")
head(year_1994)
```

##Compare Cars
```{r}
compare_cars <- function(df, queries, max_results_per_query = 1) {
  # For each query, get top match(es) using search_cars()
  matched <- purrr::map_dfr(
    queries,
    ~ search_cars(df, query = .x, max_results = max_results_per_query)
  )

  # Optional: add a unique car_id column for readability
  matched <- matched %>%
    mutate(car_id = paste(year, make, model)) %>%
    select(car_id, weekend_score, daily_score, total_score,
           mpg, hp, displacement, transmission, fuel)

  return(matched)
}

```

###Test: Compare
```{r}
compare_cars(dougscore, c("mclaren f1", "ferrari f40", "porsche 918"))
```

##Plot: Compare
```{r}
library(ggplot2)

plot_comparison <- function(df, queries, metric = "total_score", max_results = 1) {
  # Use search_cars to get top match(es) per query
  matched <- purrr::map_dfr(
    queries,
    ~ search_cars(df, query = .x, max_results = max_results)
  )

  # Create a car_id label
  matched <- matched %>%
    mutate(car_id = paste(year, make, model))

  if (!metric %in% names(matched)) {
    stop(paste("Metric", metric, "not found in data."))
  }

  ggplot(matched, aes(x = reorder(car_id, .data[[metric]]), y = .data[[metric]])) +
    geom_col(fill = "#4ad493") +
    labs(title = paste("Comparison of", metric),
         x = "Car", y = metric) +
    theme_minimal() +
    coord_flip()
}
```

###Test: Plot Compare
```{r}
plot_comparison(dougscore, queries = c("mclaren f1", "hummer ev", "nissan", "amg"), metric = "hp")
```

##Plot: Metric by Make
```{r}
plot_metric_by_make <- function(df, metric = "mpg", top_n = 10) {
  df %>%
    group_by(make) %>%
    summarize(avg_value = mean(.data[[metric]], na.rm = TRUE)) %>%
    slice_max(avg_value, n = top_n) %>%
    ggplot(aes(x = reorder(make, avg_value), y = avg_value)) +
    geom_col(fill = "#4ad493") +
    labs(title = paste("Top", top_n, "Makes by", metric),
         x = "Make", y = paste("Average", metric)) +
    coord_flip() +
    theme_minimal()
}
```

###Test: Metric by Make
```{r}
plot_metric_by_make(dougscore, metric = "hp")
```

##Plot: DougScore vs metric
```{r}
plot_score_vs_spec <- function(df, x = "mpg", y = "total_score", label_top = 5) {
  df <- df %>%
    mutate(car_id = paste(year, make, model))

  p <- ggplot(df, aes(x = .data[[x]], y = .data[[y]])) +
    geom_point(alpha = 0.6) +
    labs(title = paste(y, "vs", x), x = x, y = y) +
    theme_minimal()

  if (!is.null(label_top)) {
    top_cars <- df %>%
      slice_max(order_by = .data[[y]], n = label_top)

    p <- p + ggrepel::geom_text_repel(
      data = top_cars,
      aes(label = car_id),
      max.overlaps = Inf,
      size = 3
    )
  }

  p
}
```

###Test: Plot DougScore vs Metric
```{r}
plot_score_vs_spec(dougscore, x = "", y = "daily_score")
```
##Plot: Score by CarPlay
```{r}
plot_score_by_carplay <- function(df, score = "total_score") {
  df %>%
    filter(!is.na(carplay)) %>%
    mutate(carplay = ifelse(carplay == 1, "Has CarPlay", "No CarPlay")) %>%
    ggplot(aes(x = carplay, y = .data[[score]], fill = carplay)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.6) +
    geom_jitter(width = 0.1, alpha = 0.3) +
    labs(
      title = paste(score, "by CarPlay Availability"),
      x = NULL, y = score
    ) +
    theme_minimal() +
    theme(legend.position = "none")
}
```

###Test: Plot Score by CarPlay
```{r}
plot_score_by_carplay(dougscore, score = "features")
```
##Find Similar Cars
```{r}
find_similar_cars <- function(df, query, n = 5) {
  # Find the target car
  target <- search_metric(df, query, max_results = 1)
  if (nrow(target) == 0) stop("Car not found.")

  # Select numeric features for comparison
  compare_fields <- c("total_score", "torque", "hp", "displacement", "styling")

  df_filtered <- df %>%
    filter(if_all(all_of(compare_fields), ~ !is.na(.)))

  # Normalize features
  rec <- recipe(~ ., data = df_filtered[compare_fields]) %>%
    step_normalize(all_predictors()) %>%
    prep()
  df_norm <- bake(rec, df_filtered[compare_fields])

  # Normalize target car
  target_norm <- bake(rec, target[compare_fields]) %>% as.numeric()

  # Compute Euclidean distance
  distances <- apply(df_norm, 1, function(row) sqrt(sum((row - target_norm)^2)))

  df_filtered <- df_filtered %>%
    mutate(car_id = paste(year, make, model),
           similarity = -distances) %>%
    arrange(desc(similarity))

  df_filtered %>%
    filter(car_id != paste(target$year, target$make, target$model)) %>%
    slice_head(n = n)
}
```

###Test: Find Similar Cars
```{r}
find_similar_cars(dougscore, "Ferrari", n = 5)
```

##Get Car Summary
```{r}
get_car_summary <- function(df, query) {
  car <- search_metric(df, query, max_results = 1)
  if (nrow(car) == 0) stop("Car not found.")

  car <- car %>%
    mutate(car_id = paste(year, make, model))

  # Create a paragraph-style description
  summary_text <- glue::glue(
    "{car$car_id} scores {car$total_score}/100 on DougScore, with a {car$weekend_score}/50 for weekend driving and {car$daily_score}/50 for daily usability. ",
    "It has a {car$transmission} transmission and a {car$displacement}L engine producing {car$hp} horsepower. ",
    "Fuel economy is around {car$mpg} ", 
    if(car$fuel == "electric") "MPGe. " else "mpg. ",
    if (!is.na(car$carplay) && car$carplay == 1) "It also comes with Apple CarPlay. " else "",
    if(car$fuel != "electric") "This car is categorized as a {car$fuel} vehicle and has {car$cylinders} cylinders." else "",
    .trim = TRUE
  )

  list(
    car_data = car %>% select(car_id, weekend_score, daily_score, total_score, mpg, hp, displacement, transmission, fuel, carplay),
    summary = summary_text
  )
}

```

###Test: Get Car Summary
```{r}
get_car_summary(dougscore, "MAZDA Cx")
```
##LLM Get Car Summary
```{r}
get_car_summary_llm <- function(df, query, engine = "gpt-3.5-turbo") {
  car <- search_metric(df, query, max_results = 1)
  if (nrow(car) == 0) stop("Car not found.")

  # Format info to send to the LLM
  input_text <- glue::glue(
    "Write a short, engaging car review summary for the following specs:\n",
    "Car: {car$year} {car$make} {car$model}\n",
    "DougScore: {car$total_score}/100 (Weekend: {car$weekend_score}, Daily: {car$daily_score})\n",
    "Transmission: {car$transmission}\n",
    "Displacement: {car$displacement}L\n",
    "Horsepower: {car$hp} hp\n",
    "Fuel Type: {car$fuel_type}\n",
    "Fuel Economy: {car$mpg} mpg\n",
    if (!is.na(car$carplay) && car$carplay == 1) "Includes CarPlay.\n" else "",
    .trim = TRUE
  )

  res <- httr::POST(
    url = "https://api.openai.com/v1/chat/completions",
    httr::add_headers(
      Authorization = paste("Bearer", Sys.getenv("OPENAI_API_KEY")),
      `Content-Type` = "application/json"
    ),
    body = jsonlite::toJSON(list(
      model = engine,
      messages = list(
        list(role = "system", content = "You are a car journalist writing snappy, data-informed vehicle reviews."),
        list(role = "user", content = input_text)
      ),
      temperature = 0.7
    ), auto_unbox = TRUE)
  )

  output <- httr::content(res)
  summary_text <- output$choices[[1]]$message$content

  list(
    car_data = car %>% select(year, make, model, total_score, mpg, hp),
    summary = summary_text
  )
}
```


